---
title: "p8105_hw2_Ec3342"
author: "Ekta Chaudhary"
date: "30/09/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(dplyr)
library(knitr)
```

# Question 1

## Reading and cleaning the Mr. Trash Wheel sheet. 
#Mentioned the path to the xlsx file using path = "./data/Trash_wheet_data.xlsx", specified the sheet to read by providing the name of the sheet which is Mr. Trash Wheel here. Cleaned the variable names so that they are reasonable, omitted the rows that do not include dumpster-specific data, rounded the number of sports balls to the nearest integer.

```{r trashwheel_data}
trashwheel_data = 
  read_excel(path = "./data/Trash_wheet_data.xlsx", sheet = "Mr. Trash Wheel") %>%
  janitor::clean_names() %>%
  select(-x15, -x16, -x17) %>%
  drop_na(.,dumpster) %>%
  mutate(sports_balls = as.integer(sports_balls)
    )
```

## Reading and cleaning the precipitation data for 2017 and 2018. 
#Omitted the rows without precipitation data and added a variable year which is = 2017 for precipitaion_data_2017 and 2018 for precipitation_data_2018.Combined both the precipitation data (precipitation_data_2017 and precipitation_data_2018) to form combined_data. In the combined_data used month.name to convert the month from numeric to character variable.

```{r precipitation_data}

precipitation_data_2017 =
  read_excel(path = "./data/Trash_wheet_data.xlsx",sheet = "2017 Precipitation", skip = 1 ) %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(year = 2017)

precipitation_data_2018 =
  read_excel(path = "./data/Trash_wheet_data.xlsx",sheet = "2018 Precipitation",skip = 1) %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(year = 2018)

combined_data = union_all(precipitation_data_2017, precipitation_data_2018) %>%
  mutate(month = month.name[month])
```
## Describing both the datasets!

*The trashwheel_data comprises of `r nrow(trashwheel_data)` observations of `r ncol(trashwheel_data)` variables.

*The precipitation_data_2017 comprises of `r nrow(precipitation_data_2017)` observations of `r ncol(precipitation_data_2017)` variables.

*The precipitation_data_2018 comprises of `r nrow(precipitation_data_2018)` observations of `r ncol(precipitation_data_2018)` variables.

*The number of observations in combined_data is `r nrow(combined_data)` and the number of columns are `r ncol(combined_data)`.

*The key variables are dumpster, month, year.

*The total precipitation in 2018 was `r sum(pull(precipitation_data_2018, total))`.

*The median number of sports ball in a dumpster in 2017 was `r trashwheel_data %>% select(sports_balls, year) %>% filter(year == 2017) %>% summarise(median = median(sports_balls))`.

# Question 2

##First read the dataset pols-month.csv and cleaned the data in pols-month.csv. Used separate() to break up the variable mon into integer variables year, month, and day; replaced month number with month name and converted it to lower case to maintain consistency with rest of the datasets. Year is converted into an integer. Removed the variable day. Created a president variable taking values gop and dem, and removed prez_dem and prez_gop.

```{r pols_data}
pols_data = read_csv(file = "./data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(col = mon, into = c("year", "month", "day")) %>% 
  select(prez_gop, prez_dem, everything()) %>%
  pivot_longer(
    prez_gop:prez_dem, names_to = "president", names_prefix = "prez_") %>%
  mutate(
   month = as.numeric(month),
   month = month.abb[month],
   month = str_to_lower(month),
   year  = as.integer(year)
  ) %>%
  select(-day)
```

##Cleaning the data in snp.csv using a similar process to the above. Seperated the variable mon into year,month and day using separate().Replaced month number with month name and converted it to lower case to maintain consistency with rest of the datasets.Year is converted into an integer. Removed the variable day. For consistency across datasets, arranging according to year and month using the arrange() function, and organizing so that year and month are the leading columns using the select() function.

```{r snp_data}

snp_data = read_csv(file = "./data/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(col = date, into = c("month", "day", "year")) %>% 
  arrange(as.integer(year), as.integer(month)) %>%
  select(year, month, close) %>%
  mutate(
    month = as.numeric(month),
    month = month.abb[month],
    month = str_to_lower(month),
    year = as.integer(year)
    ) 
```
##Tidying the unemployment data so that it can be merged with the previous datasets. Using pivot_longer to go from the wide to long format.Organizing so that year and month are the leading columns using the select() function. Year is converted into an integer. So, in all the three datasets month is given by name in lower case, year is an integer, day is removed and year,month are the leading columns.  

```{r unemployment_data}

unemployment_data =
  read_csv(file = "./data/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec, names_to = "month", values_to = "unemployment_rate") %>%
  select(year, month, unemployment_rate) %>%
  mutate(year = as.integer(year))
```

##merged_data is formed by joining snp into pols using year and month as keys. final_merged_data is formed by joining the merged_data with unemployment_data using year and month as keys.

```{r merged_data}

merged_data = left_join(snp_data, pols_data, by = c("year","month"))

final_merged_data = left_join(merged_data, unemployment_data, by = c("year","month"))
```
##Description of the datasets

* The dataset "pols_data" contains `r nrow(pols_data)` observations of `r ncol(pols_data)` variables related to the number of national politicians who are democratic or republican at any given time.

* The dataset “snp_data” contains `r nrow(snp_data)` observations of `r ncol(snp_data)` variables related to Standard & Poor’s stock market index (S&P), often used as a representative measure of stock market as a whole. It has the date (year, month) of the observation and the closing values of the S&P stock index on the associated date.

* The dataset "unemployment_data" contains the percentage of unemployment (unemployment rate) for each month from 1948-2015. This dataset contains `r nrow(unemployment_data) ` observations and `r ncol(unemployment_data)` columns.

* The final_merged_data contains `r nrow(final_merged_data)` observations of `r ncol(final_merged_data)` variables. This final_merged_data is formed by joining the merged_data which has `r nrow(merged_data)` rows and `r ncol(merged_data)` columns and the unemployment_data which has `r nrow(unemployment_data)` rows and `r ncol(unemployment_data)` columns. The merged_data was formed by combining pols_data (which has `r nrow(pols_data)` rows and `r ncol(pols_data)` columns) and snp_data which has `r nrow(snp_data)` and `r ncol(snp_data)` columns. The final_merged_data shows us the closing values of the S&P stock index, number of national politicians who are democratic or republican and the percentage of unemployment (unemployment rate) at any given time.

* The key variables are year, month, close, president, value, unemployment rate.

# Question 3

##Loading and tidying the data Popular_Baby_Names and creating a dataset called pop_baby. Changing the case structure of gender, ethnicity and childs first name to lower case using mutate() function to maintain uniformity throughout. Removing all the duplicate rows using distinct().

```{r popular baby}

pop_baby = read_csv(file = "./data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(
    gender = str_to_lower(gender), ethnicity = str_to_lower(ethnicity),
    childs_first_name = str_to_lower(childs_first_name)
    ) %>%
  distinct()
```
##Producing a well-structured, reader-friendly table using kable() function showing the rank in popularity of the name “Olivia” as a female baby name over time; The table shows ethnicities in the rows and year in the columns. 

```{r olivia}

olivia = filter(pop_baby, gender == "female", childs_first_name == "olivia") %>%
  select(-gender, -childs_first_name, -count)
olivia[olivia == "asian and paci"] <- "asian and pacific islander"
olivia[olivia == "black non hisp"] <- "black non hispanic"
olivia[olivia == "white non hisp"] <- "white non hispanic"
olivia = pivot_wider(olivia, names_from = "year_of_birth", values_from = "rank") 
kable(olivia, format = "html")
```
##Producing a similar table called pop_boy showing the most popular name among male children over time for male

```{r pop_boy}

pop_boy = filter(pop_baby, gender == "male", rank == 1) %>%
  select(-gender, -count, -rank)
pop_boy[pop_boy == "asian and paci"] <- "asian and pacific islander"
pop_boy[pop_boy == "black non hisp"] <- "black non hispanic"
pop_boy[pop_boy == "white non hisp"] <- "white non hispanic"
pop_boy = pivot_wider(pop_boy, names_from = "year_of_birth", values_from = "childs_first_name") 
kable(pop_boy, format = "html")
```
##Producing a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis) for male, white non-hispanic children born in 2016.

```{r}

filter(pop_baby, gender == "male", ethnicity == "white non hispanic", year_of_birth == 2016) %>%
  ggplot(aes(x = rank,y = count)) + geom_point()

```



